---
title: "Week 3 homework assignment"
author: Jay Kim
date: Due Date
output: github_document
editor_options:
  chunk_output_type: inline
---


## Part 1: Working with Midwest Data
```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
View(midwest)
```


```{r}
g1 <- g1 + scale_y_continuous(breaks=seq(0, 1000000, 200000), labels = function(x){paste0(x/1000, 'K')})
```
This makes a graph called "g1". The graph has a y axis from 0 to 1,000,000 and 200,000 is the intervals. But when the graph is outputted, each of the intervals is divided by 1,000 so it is more visually appeasing, but also attaches a 'K' at the end of each number, so it indicates that it is, for example, '200K' not '200'. 


# Part 2
Exploring the relationship of population totals by state. Included a clear title, and change the xlab and ylab to be easy to read words (labels). 

```{r}
midwest%>%
  group_by(state)%>%
  summarize(popbystate= sum(poptotal))%>%
  select(state, popbystate)%>%
  ggplot(mapping = aes(x=state, y=popbystate)) +
  geom_col()

```

3. Made a histogram and density plot for the percent of people below poverty

```{r}
#answer
midwest%>%
  select(percbelowpoverty)%>%
  ggplot(mapping = aes(x=percbelowpoverty))+
  geom_density()+
  geom_histogram()

midwest%>%
  select(percbelowpoverty)%>%
  ggplot(mapping = aes(x=percbelowpoverty))+
  geom_density(fill="blue")
```

## Part 3

```{r}
# scatterplot, student code:
gg_new <- ggplot(metro_nash_schools, aes(x=Economically.Disadvantaged,y=Limited.English.Proficiency))+
  geom_point(aes(col=School.Level), size =2) +
  labs(
    title="Proficiency vs Economically Disadvantaged",
    subtitle = "Metro Nash School",
    y= "Eng Proficiency",
    x= "Econmonically Disadvanged",
    color= "school level"
  )+
  theme_classic()
gg_new

```

I thought this graph was an effective graph for multiple reasons. First, it doesn't contain too much information, as in it sticks to graphing 2 variables and one can see the relationship between them. Second, the colors make it easy to distinguish what is going on and then the axis are clearly labeled. The scatter plot is also the ideal way to display the potential relationship between the variables English Proficiency and Economically Disadvantaged. 
 









