---
title: "Week 4 homework assignment"
author: Student Name
date: Due Date
output: github_document
editor_options:
  chunk_output_type: inline
---

## DSI-EDA
## Professor Cassy Dorff
```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
```
#Graph 1
```{r}
ggplot(week4_hw_data, aes(x=trump, y=discrim))+
  geom_point(col="blue",size = 1.5)+
  scale_x_continuous(breaks = seq(0, 1, 0.1), labels= function(x){paste0(x*100, '%')})+
  scale_y_continuous(breaks = seq(0, 1, 0.1), labels= function(x){paste0(x*100, '%')})+ #how to do this correctly
  labs(x = "2016 Trump vote",
       y = "% saying discrimination")+
  #geom_text(aes(label= ifelse(trump<.2, as.character(state), ''),),hjust=1.2, vjust =0)+
  #geom_text(aes(label= ifelse(discrim<.2, as.character(state), '')), hjust=1.2, vjust =0)+
  geom_abline(slope=-0.8, intercept = 0.9)+
  theme_bw()
```
Yes, if you were to take out the leftmost and right/bottom most outlier and made the x and y axis of the same length, the graph would have little to no downward trend. I do think a relation exists, just maybe not as dramatic as the graph may portray. To get a more accurate relationship, we could make the graph axis of equal length.

#Graph 2
```{r}
week4_hw_data$party <- ifelse(week4_hw_data$state %in% c("CA", "NY","WA", "OR","NV","CO","NM","MN","IL","VA","MA", "RI", "CT", "NJ", "DE", "DC", "VT", "NH" ), 'Biden',      
                       ifelse(week4_hw_data$state %in% c("ID", "MT", "WY", "UT", "ND", "SD", "KS", "OK", "TX", "AR", "LA", "MS", "AL", "TN", "KY", "IN", "WY" ), 'Trump', "Toss Up"))
week4_hw_data %>% 
  ggplot(aes(x= trump, y= discrim, col= party))+
  geom_point()+
  geom_text(aes(label= ifelse(trump<1, as.character(state), ''),),hjust=1.5, vjust =0, size =2)+
  scale_colour_manual(values =c("blue", "black","red"))
  
```
With this year's election so close, in this graph I have specified which states are leaning towards electing which president based on the recent polls. While the results are unsurprising, it is interesting to see visually what is going happening What is also unsurprising is that the states leaning Trump tend to think discrimination is less prevalent, and the states leaning Biden seem to think the it is more prevalent. There could be a few explanations to why this may be, but that could be a conversation for another time.  


#Graph 3
Previously, we explored data about wine. 
```{r}
library(readr)
Wine <- read_delim("~/Documents/EDA/eda20-live-code/mod4/winequality-red.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)
```


#Graph 4
```{r}
ggplot(winequality_red, aes(x=quality, y= alcohol)) +
      geom_bar(stat='identity') +
      theme_bw()
```
A bar plot with the wine data to explore our original question about the relationship between alcohol content and quality. 

#Graph 5
```{r}
ggplot(winequality_red, aes(x=quality, y= alcohol)) +
      geom_jitter(aes(col=`fixed acidity`)) +
      scale_color_viridis(option = "D", direction = -1, alpha = .6) +
      labs(title="Alcohol Content, Fixed Acidity and Wine Quality", y="Alcohol Content", x="Wine Quality")  +
      labs(color='Fixed Acidity') +
      theme_bw()+
  theme(legend.position = "top")
```
One thing we can see from this graph is that it seems like a lot of the high content of wine have a fixed acidity of 5!

