---
title: "Analysis on Life Expectancy"
author: "Jay Kim"
output: 
  html_document:
    code_folding: hide
--- 

# Part One: Introduction
This data set contains a lot of information on life expectancy in many many different countries across the globe. The World Health Organization, more commonly referred to as 'WHO', began procuring annual life tables for all Member States in 1999. These life tables are a basic input to all WHO estimates of global, regional and country-level patterns and trends in all-cause and cause-specific mortality. The WHO then published these life expectancy estimates for 194 countries total between the years of 2000 and 2015. The WHO seemed to have collected this data for a few reasons. They wanted to see any chances in mortality rates within different countries and how it changed over the course of several years- if at all. They wanted to see if gender played a role in life expectancy or current event during certain times, like the AIDS epidemic. 

```{r message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(janitor)
library(maps)
library(ggplot2)
library(ggalt)
theme_set(theme_classic())
world = map_data("world")
```
```{r cache=TRUE,message=FALSE, warning=FALSE}
life <- read_csv("life_data.csv") %>% #Adjusting variable names in data set
clean_names()
life
```

# Graph #1
This graph first plots the 10 countries where there was the greatest increase in life expectancy and the the 10 countries with the smallest increase of life expectancy(including any decreases as well) from the years 2000 to 2016. But before we can do so, we need to create a dataset that contains the a column with the value of the difference between the life ex. from the time interval 2016 and 2000. We will then we will select the necessary 20 rows. 
```{r}
new_life<- life%>%
  mutate(increase=both_sexes_2016 - both_sexes_2000)%>%
  select(country, increase, both_sexes_2016, both_sexes_2000)%>%
  arrange(desc(increase))%>%
  slice(tail(row_number(),(10)), head(row_number(),10))
new_life
```

```{r}
ggplot(new_life, aes(y=country, x = both_sexes_2000, xend = both_sexes_2016)) + 
        geom_dumbbell(color="#a3c4dc", 
                      size=0.75, 
                      point.colour.l="#0e668b") +
  labs(title = "Increase in Life Expectancy", 
       x = "Years",
       y = "Country")
geom_segment(data = life, aes(y=concerned, yend=concerned, x=20, xend=100), color="#b2b2b2", size=0.15)
```
We see that all countries saw an increase but one (if you look at the last entry), the Syrain Arab Republic. They had a decrease in life ex. by 7.3 years. 


# Graph 2
First, we had to do some data manipulation. We computed the mean and then created a new column with the difference in life expectancy from 2016 and the mean, we called the column "average".
```{r}
diff_life <- life %>%  #Adding a new column that contains the country's deviation in life expectancy from the global average
  mutate(average = both_sexes_2016 - mean(both_sexes_2016))%>%
  select(country, both_sexes_2016, average)
diff_life
```
The average life exptectancy is computed below.
```{r}
life %>%
  summarise(mean(both_sexes_2016)) #Global mean of life expectancy
```

```{r}
df <- sample_n(diff_life, size = 35)
df
ggplot(df, 
       aes(x = reorder(country, average ), y = average,
           fill = average >0))+
  geom_bar(stat = "identity")+
  coord_flip()+
  labs(x = "Countries", y="Years",
       title = "Above or Below Global Average?",
       subtitle = "The average life expentancy in 2016 was 63.22 years ")+
  scale_fill_discrete(name = "", labels = c("Below Average", "Above Avergae"))
```
We can see a few things here. This graph shows how 35 randomly selected countries' life expectancy compare to the global average: 63.22 years. If a country's expectancy was 67 for example, then it was show that that it was 4 years above average.Countries in Europe and North America have much higher life expectancy on average than countries in Southeast Asia and parts of Africa. This could be a result in the differences of living in a more developed country versus a developing country. Developing countries, such as those in Sub-Saharan Africa, deal with issues that places such as Switzerland, Norway, and Italy do not: famine, low level of education, availability of clean water, among many others. Developed countries not only do not have to worry as much about issues like those, but also have improved and robust medical care in place in case their citizens fall ill. 


# Graph 3
```{r}
library(ggplot2)
library(scales)
theme_set(theme_classic())
```

In the code below we want to see the difference in life expectancy between women and men in each country and then arrange them in descending order. If the difference (stored in 'diff') is positive, then it means that in that country, women have a higher life expectancy.
```{r}
life3 <- life%>%
  select(male_2016, female_2016, country)%>%
  mutate(diff = female_2016 - male_2016)%>%
  arrange(desc(diff))
life3
```
Russian Federation, Lithuania, and Belarus has the highest difference of about 8 years- which is a big difference! We can then look at the last few entries of this data set which will tell us the countries that have the biggest gaps but with men with the high life expectancy. We will graph the top 3 from both ends of the dataset below to analyze. But quickly, we will create a dataset that contains these 6 values.
```{r}
Big_Diff <- life3%>%
  slice(tail(row_number(),3), head(row_number(),4))
Big_Diff
```


```{r}
#how to add color by country
#How to hide code
#Is adding 1 graph enough?
Big_Diff
p <- ggplot(Big_Diff, aes(x=1, xend=2, y= male_2016, yend=female_2016, colour=country)) + 
  geom_segment(size=.75, show.legend=T) +
  geom_vline(xintercept=1, linetype="dashed", size=.1) +
  geom_vline(xintercept=2, linetype="dashed", size=.1) +
  labs(x="", y="Average Life Expectancy") +
  xlim(0.75, 2.25) + ylim(60,71)+   # X and Y axis limits 
 geom_text(label="Men", x=1, y=71, color="Black")+
  geom_text(label="Women", x=2, y=71, color="Black")
p

```
Interestingly, the top countries where the women live longer live a lot longer than the top countries where men live longer. We may have only graphed 3 countries from both criteria, but if you were to go back and look at the life3 data set, most of the values in the 'diff' column are indeed positive, meaning the women live longer.   


```{r}
life3%>%
  summarise(mean(diff))
```

Just out of curiosity I calculated the average difference between all countries, and it comes out to be 3.26 years. 


